% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\author{}
\date{}

\begin{document}

\hypertarget{technical-documentation}{%
\section{Technical Documentation}\label{technical-documentation}}

The language corpus of spoken performances by non-native speakers of
Czech, focused on the A2 language level (according to the CEFR),
required for obtaining permanent residency in the Czech Republic, is the
result of a project implemented at the Institute of Formal and Applied
Linguistics of the Faculty of Mathematics and Physics, Charles
University. The corpus contains recordings capturing the oral part of
the
\href{https://ujop.cuni.cz/UJOPEN-70.html?ujopcmsid=12:czech-language-certificate-exam-cce}{Czech
Language Certificate Exam} at the A2 level. The recordings include
dialogues between the examiner (a native speaker) and the candidate (a
non-native speaker). We have provided transcriptions of the recordings,
enriched with extensive linguistic annotations. Some recordings are
accompanied by multiple transcriptions from different annotators,
allowing for comparisons of various transcriptions of the same recording
and the assessment of the degree of agreement when converting spoken
language into written text.

The corpus is published as a specialized public database and is freely
accessible to the general public, the scientific community, educators,
and students. The database is integrated into the TEITOK system, managed
on the \href{https://lindat.cz/}{LINDAT/CLARIAH-CZ} platform.

\hypertarget{teitok}{%
\subsection{TEITOK}\label{teitok}}

\href{http://teitok.corpuswiki.org/}{TEITOK} is a framework for
creating, managing, and publishing annotated corpora. Its web interface
is implemented using a combination of PHP and JavaScript. For our
project, which combines recordings of spoken speech and their
transcriptions, the key functionality of the TEITOK environment allows
us to
\href{http://www.teitok.org/index.php?action=help\&id=wavesurfer}{create,
display, and edit recordings' transcriptions}. To work with the
recordings themselves, TEITOK utilizes the JavaScript library
\href{http://wavesurfer-js.org/}{wavesurfer}.

\hypertarget{data-storage}{%
\subsubsection{Data Storage}\label{data-storage}}

The corpus data is primarily stored in the TEITOK environment in the
form of files. In this case, the recordings are in MP3 format, while the
main components are TEITOK format files, which contain all
transcriptions and annotations, including metadata. These files are
interconnected with the corresponding recordings.

\hypertarget{structure-of-teitok-files}{%
\subsubsection{Structure of TEITOK
Files}\label{structure-of-teitok-files}}

The TEITOK format is an XML format that fully complies with the
\href{https://www.tei-c.org/}{Text Encoding Initiative (TEI)} standards,
but with a slightly different approach to tokenization. The structure of
TEITOK files in our database is as follows:

\hypertarget{header-with-metadata-teiheader}{%
\paragraph{\texorpdfstring{Header with Metadata
\texttt{\textless{}teiHeader\textgreater{}}}{Header with Metadata \textless teiHeader\textgreater{}}}\label{header-with-metadata-teiheader}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{\texttt{\textless{}fileDesc\textgreater{}}} -- File
  description

  \begin{itemize}
  \tightlist
  \item
    \textbf{\texttt{\textless{}titleStmt\textgreater{}}}: Contains the
    title of the file and information about authors and annotators.
  \item
    \textbf{\texttt{\textless{}editionStmt\textgreater{}}}: Contains
    version number.
  \item
    \textbf{\texttt{\textless{}publicationStmt\textgreater{}}}:
    Publication details, such as publisher, release date, and license.
  \item
    \textbf{\texttt{\textless{}sourceDesc\textgreater{}}}: Description
    of the source recording and a link to it.
  \end{itemize}
\item
  \textbf{\texttt{\textless{}encodingDesc\textgreater{}}} -- Description
  of encoding

  \begin{itemize}
  \tightlist
  \item
    \textbf{\texttt{\textless{}projectDesc\textgreater{}}}: A brief
    description of the project under which the data was created.
  \item
    \textbf{\texttt{\textless{}annotationDecl\textgreater{}}}: Details
    of the individual annotation steps (primary, revision, linguistic
    annotation).
  \end{itemize}
\item
  \textbf{\texttt{\textless{}profileDesc\textgreater{}}} -- Profile of
  the text

  \begin{itemize}
  \tightlist
  \item
    \textbf{\texttt{\textless{}langUsage\textgreater{}}}: Language used
    (Czech).
  \item
    \textbf{\texttt{\textless{}textClass\textgreater{}}}: Document
    metadata:

    \begin{itemize}
    \tightlist
    \item
      \texttt{database}: Database name.
    \item
      \texttt{exam-id}: Exam identifier.
    \item
      \texttt{cefr-level}: CEFR level. This database contains recordings
      exclusively from A2 level exams.
    \item
      \texttt{task-number}: Task number.
    \item
      \texttt{preannot-source}: Source of preliminary annotation.
    \item
      \texttt{annotator}: Annotator code.
    \item
      \texttt{canonical}: A value of \texttt{1} indicates a canonical
      transcription.
    \end{itemize}
  \end{itemize}
\end{enumerate}

\hypertarget{main-content-text}{%
\paragraph{\texorpdfstring{Main Content
\texttt{\textless{}text\textgreater{}}}{Main Content \textless text\textgreater{}}}\label{main-content-text}}

The \texttt{\textless{}text\textgreater{}} section contains individual
segments of spoken speech structured using
\texttt{\textless{}u\textgreater{}} elements: -
\textbf{\texttt{\textless{}u\textgreater{}}}: Each
\texttt{\textless{}u\textgreater{}} element represents a segment of
speech and has attributes: - \texttt{start} and \texttt{end}: Start and
end time in seconds. - \texttt{who}: Speaker (e.g., ``EXAM\_1'' for the
examiner and ``CAND\_1'' for the candidate). -
\textbf{\texttt{\textless{}s\textgreater{}}}: Each sentence is marked
with the \texttt{\textless{}s\textgreater{}} element. -
\textbf{\texttt{\textless{}tok\textgreater{}}}: Token elements whose
attributes describe lemma, part of speech, morphological features, and
syntactic relations. - \textbf{\texttt{\textless{}anon/\textgreater{}}}:
Anonymized segment of the recording. -
\textbf{\texttt{\textless{}gap\ reason="unintelligible"/\textgreater{}}}:
Unintelligible segment of the recording.

\hypertarget{preparation-of-teitok-files}{%
\subsubsection{Preparation of TEITOK
Files}\label{preparation-of-teitok-files}}

The preparation of TEITOK files took place in several phases:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Preliminary Annotation}. In the research associated with the
  creation of the database, we compared direct manual annotation with
  manual post-editing of outputs from automatic speech recognition
  systems. Thus, manual annotation may be based on automatically
  prepared preliminary annotation. The source of the preliminary
  annotation is distinguished using the \texttt{preannot-source}
  attribute, which can have the following values:

  \begin{itemize}
  \tightlist
  \item
    \texttt{from\_scratch}: Completely manual annotation, i.e., the
    preliminary annotation is empty.
  \item
    \texttt{from\_whisperX}: Preliminary annotation obtained using the
    \href{https://github.com/m-bain/whisperX}{WhisperX} system.
  \item
    \texttt{from\_mixed}: Preliminary annotation obtained by randomly
    combining outputs from four systems at the level of utterances. When
    the preliminary annotation was not empty, we converted it into the
    basic version of the TEITOK format. At the end of this phase, the
    transcriptions contained segments divided into utterances (the
    \texttt{\textless{}u\textgreater{}} elements), assignment of
    speakers to utterances (the \texttt{who} attribute), and time
    alignment with the recording (the \texttt{start} and \texttt{end}
    attributes).
  \end{itemize}
\item
  \textbf{Manual Annotation}. After uploading the files, trained
  annotators performed manual annotation in the TEITOK environment,
  during which they created or corrected transcriptions, assigned
  speakers to utterances, and aligned utterances with the recording
  using timestamps. The recordings were anonymized in accordance with
  the requirements of the Institute for Language and Preparatory Studies
  of Charles University (ILPS CU), which provided the audio recordings
  for the corpus. Some annotators, out of caution, anonymized even data
  that did not need to be anonymized (e.g., fictitious names).
\item
  \textbf{Revision}. Manual review of the manual annotations by a
  co-author of the database.
\item
  \textbf{Normalization}. Automatic adjustment of transcriptions that
  removes discrepancies in speaker names, orders utterances according to
  start time, and assigns new sequential IDs to utterances.
\item
  \textbf{Segmentation by Tasks and Selection}. The provider of the
  recordings (ILPS CU) permitted the publication of only selected tasks.
  We had to cut these from the recordings and adjust timestamps in the
  transcriptions to preserve the alignment of utterances in the
  transcription with the recording. We used the
  \href{https://www.ffmpeg.org/}{FFmpeg} tool for cutting the
  recordings.
\item
  \textbf{Linguistic Annotation}. Until this phase, the utterances in
  the transcriptions had not been further structured. In this phase, we
  divided the text into sentences (the
  \texttt{\textless{}s\textgreater{}} element) and then into tokens (the
  \texttt{\textless{}tok\textgreater{}} elements). At the token level,
  the transcriptions are automatically linguistically annotated. Each
  token is assigned a lemma (the \texttt{lemma} attribute),
  language-specific morphological tag (the \texttt{xpos} attribute),
  part of speech, and morphological properties according to the
  categorization of the
  \href{https://universaldependencies.org/}{Universal Dependencies}
  project (the \texttt{upos} and \texttt{feats} attributes).
  Additionally, each token is assigned a reference to the parent ID
  according to dependency syntax rules (the \texttt{head} attribute) and
  the type of dependency of the token in relation to its parent (the
  \texttt{deprel} attribute). For linguistic annotation, including
  tokenization, we used the
  \href{https://ufal.mff.cuni.cz/udpipe/2}{UDPipe 2} tool, specifically
  the model \texttt{czech-pdt-ud-2.12-230717} for Czech. Although it is
  possible to perform tokenization and automatic linguistic annotation
  directly in the TEITOK environment, we carried out this process
  separately. The reason is that the tokenization method in the TEITOK
  environment differs from the one optimized for UDPipe, which could
  lead to errors when combining these two steps.
\item
  \textbf{Completion of the TEI Header}. Finally, we supplemented the
  header according to all available metadata to comply with TEI
  standards.
\end{enumerate}

All tools and scripts (primarily in Python 3 and BASH) are available in
the \href{https://github.com/ufal/evaldio}{public repository of the
project} in the \texttt{data\_preparation} directory.

\hypertarget{querying-searching-and-filtering}{%
\subsubsection{Querying, Searching, and
Filtering}\label{querying-searching-and-filtering}}

Rapid querying, searching, and filtering are enabled by the integrated
\href{https://cwb.sourceforge.io/files/CQP_Manual.pdf}{CQP Query
Processor}, a key component of the
\href{https://cwb.sourceforge.io/}{IMS Open Corpus Workbench (CWB)}
toolkit. CQP converts XML-formatted corpora into binary format and
efficiently indexes them. Querying in indexed corpora is conducted using
the
\href{https://www.cambridge.org/sketch/help/userguides/CQL\%20Help\%201.3.pdf}{CQL}
language, which is a standard in corpus linguistics. TEITOK also offers
a Query Builder, in which users can specify a query by filling out a
form. The results of the query returned from CQP are subsequently
processed using TEITOK and presented to the user in a clear format.
Query results can be downloaded in XML format.

\end{document}
